Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme

### 2020-04-25 _region_.tex
Cambios 24-4-2020

\message{ !name(LME-MPM.tex)}% %%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
% %
% % This is a general template file for the LaTeX package SVJour
% % for Springer journals.          Springer Heidelberg 2010/09/16
% %
% % Copy it to a new file with a new name and use it as the basis
% % for your article. Delete % signs as needed.
% %
% % This template includes a few options for different layouts and
% % content for various journals. Please consult a previous issue of
% % your journal as needed.
% %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %
% % First comes an example EPS file -- just ignore it and
% % proceed on the \documentclass line
% % your LaTeX will extract the file if required
% \begin{filecontents*}{example.eps}
% %!PS-Adobe-3.0 EPSF-3.0
% %%BoundingBox: 19 19 221 221
% %%CreationDate: Mon Sep 29 1997
% %%Creator: programmed by hand (JK)
% %%EndComments

% gsave
% newpath
%   20 20 moveto
%   20 220 lineto
%   220 220 lineto
%   220 20 lineto
% closepath
% 2 setlinewidth
% gsave
%   .4 setgray fill
% grestore
% stroke
% grestore
% \end{filecontents*}
% %
% \RequirePackage{fix-cm}
% %
% %\documentclass{svjour3}                     % onecolumn (standard format)
% %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
% %\documentclass[smallextended]{svjour3}       % onecolumn (second format)
% \documentclass[twocolumn]{svjour3}          % twocolumn
% %
% % flush right qed marks, e.g. at end of proof
% \smartqed

% \usepackage[lofdepth,lotdepth]{subfig}
% \usepackage{graphicx}
% \usepackage{subcaption}

% \usepackage{amsmath,amsthm,amssymb}

% \usepackage{amsfonts}
% \usepackage{epstopdf}
% \usepackage{mathrsfs} % para formato de letra
% \usepackage{hyperref}
% \usepackage{enumitem}
% \usepackage{array}
% \usepackage{tabularx}
% \usepackage{supertabular}
% \usepackage{fancyhdr}
% \usepackage{multirow}
% \usepackage{color}
% \usepackage{makeidx}
% \usepackage{setspace}
% \usepackage{epsfig}

% \usepackage{pgf}
% % \usepgfplotslibrary{external} 
% % \tikzexternalize

% %\usepackage{latexsym}

\documentclass[preprint,12pt,a4paper]{elsarticle}

@@

@@

\begin{abstract}
  Material Point Method (MPM) has arisen in the recent years as an
  alternative to Finite Element Method (FEM) under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard MPM time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time integration scheme has been proposed. Its superior performance mitigates the 
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in MPM 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) a robust substitute for the wide range of shape
  function in MPM. Local \textit{max-ent} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

\begin{keyword}
  LME \sep MPM \sep Explicit predictor-corrector \sep Dynamic problems}
\end{keyword}

@@

\section{Introduction}
\label{intro}
Since the proposal of MPM by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the Fluid
Implicit Particle (FLIP) method~\cite{Brackbill1986}. It popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the generalized interpolation material point method
(GIMP) Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
MPM proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to MPM framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

This document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) as a robust substitute for the wide range of shape
  function in MPM. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite element method (FEM) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions~\cite{Arroyo2006}. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)~\cite{Navas2018a} and Li {\it et al.} (2012)~\cite{Li2012} for
Optimal Transportation Meshfree (OTM) method. And more
recently under MPM framework by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020} but without exploring the benefits of
the regularization parameter $\beta$.

This techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using
techniques~\cite{Tran2019e}. These numerical inaccuracies also known
as wiggles are due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises is to add a
nonphysical damping source to the equilibrium equations, this
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing the accuracy of the time integration
scheme. One of the most popular is the implicit GIMP (iGIMP)
Charlton {\it et al.} (2017)~\cite{Charlton2017}, more recently Tran \&
Solowski (2019)~\cite{Tran2019e} proposed a generalised-$\alpha$ scheme
for MPM with promising results but at the expense of increasing
the computational effort. In this paper a less time consuming and high
efficient explicit predictor corrector integration method has been
proposed. It consists in an accommodation of the Newmark
predictor-corrector or central difference explicit (CD). We have
choose this method among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because it simplicity and it good
performance dealing with solids dynamics problems under a mesh-free
framework in~\cite{Navas2018a}.

@@

The article is organized as follows. Section \ref{sec:Notation}
briefly reviews the notation here employed. Next, Section \ref{sec:derivation-mpm}
is devoted to present briefly the governing equations of the elastic
problem, the variational formulation and the Galerkin procedure. In Section \ref{sec:epc-algor-mpm} an explicit predictor-corrector
time integration scheme for MPM is proposed. Section
\ref{sec:local-max-ent} briefly reviews the local \textit{max-ent}
basis functions here employed. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are presented. Finally,
conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{Brief note concerning the notation}
\label{sec:Notation}
In what follows, we will adopt the following convention. All the
physical variables involved in this problem are collected in Table
\ref{tab:notation-table}. Three kind of subscript or superscript are used within paper. The subscript$\qed_p$ is used to define a particle variable. While the
subscript$\qed_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\qed^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\qed}$ and
$\ddot{\qed}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $(\cdot)$ and $(\colon)$ means the first and second contraction of a tensor, $\Div{\qed}$ denotes the
divergence operator, and finally $\Grad{\qed}$ and $\GradS{\qed}$
denotes the gradient and its symmetric part. Einstain subscripts
convention is adopted therefore repeated index means addition. 
\begin{table}
  \centering
  \caption{Physical variables involved in the problem}
  \label{tab:notation-table}
  \begin{tabular}{c c c}
    $\rho$ & Density field & Scalar \\
    $\vect{a}$ & Acceleration field & First order tensor \\
    $\vect{v}$ & Velocity field & First order tensor \\
    $\vect{u}$ & Displacement field & First order tensor \\
    $\vect{x}$ & Global coordinates & First order tensor \\
    $\vect{\xi}$ & Local coordinates & First order tensor \\
    $\tens{\sigma}$ & Cauchy stress tensor & Second order tensor \\
    $\tens{\varepsilon}$ & Cauchy strain tensor & Second order tensor \\
    $\tens{D}$ & Constitutive tensor & Fourth order tensor\\
  \end{tabular}
\end{table}

\section{Derivation of MPM procedure}
\label{sec:derivation-mpm}
The aim of this section is to provide an overview of the standard
explicit MPM algorithm~\cite{Sulsky1994}. In consequence it
is structured as follows: the governing equations will be introduced
in \ref{sec:The boundary-value-problem}, later the variational statement of
the problem will be presented \ref{sec:variational-formulation}, and
finally a discretization procedure thorough Galerkin is performed in
\ref{sec:Galerkin-procedure}. The method has three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation thorough a FEM-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence MPM can be regarded as a
Lagrangian-Eulerian method where particles carries on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation.

\subsection{The boundary-value problem}
\label{sec:The boundary-value-problem}
In MPM the continuum mechanics approach is considered. So on, let
define a continuum domain $\varOmega$ occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and $\partial
\varOmega$ the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field $\vect{u}$ allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field $\vect{u}$ through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from $\vect{u}$ the strain field is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
$\tens{\sigma}$, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to  $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

\subsection{Variational formulation}
\label{sec:variational-formulation}
To write the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative} represents the weak form
formulation of the problem.

\subsection{Galerkin procedure}
\label{sec:Galerkin-procedure}

@@

FEM, in MPM a double discretization procedure is performed as we will

@@

\section{Explicit predictor-corrector scheme for MPM.}
\label{sec:epc-algor-mpm}

>>>>

\message{ !name(LME-MPM.tex)}
\documentclass[preprint,12pt,a4paper]{elsarticle}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init glossaries
\usepackage[acronym]{glossaries}

\makeglossaries

\newglossaryentry{domain}
{
  name=$\varOmega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\varOmega$,
  description={Boundary of the continuum domain $\varOmega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{rho}
{
  name=$\rho$,
  description={Describes the scalar density field}
}

\newglossaryentry{a}
{
  name=$\vect{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vect{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vect{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vect{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vect{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

\newglossaryentry{stress}
{
  name=$\tens{\sigma}$,
  description={Second order tensor which means the Cauchy stress tensor}
}

\newglossaryentry{strain}
{
  name=$\tens{\varepsilon}$,
  description={Second order tensor which means the Cauchy strain tensor}
}

\newglossaryentry{Constitutive}
{
  name=$\tens{D}$,
  description={Four order tensor which means the constitutive response
    of the material}
}

\newglossaryentry{LME-beta}
{
  name=$\beta$,
  description={Regularization or thermalization parameter of the
    LME$_{\beta}$ Pareto set}
}

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}

% End glossaries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{MPM} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time
  integration scheme has been proposed. Its superior performance
  mitigates the presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in \acrshort{MPM} 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \acrfull{lme} approximation schemes a
  robust substitute for the wide range of shape function in
  \acrshort{MPM}. Local \textit{max-ent} approximation may be regarded
  as a \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

\begin{keyword}
  \acrshort{lme} \sep \acrshort{mpm} \sep Explicit predictor-corrector \sep Dynamic problems}
\end{keyword}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}. It popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to \acrshort{mpm} framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

This document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) as a robust substitute for the wide range of shape
  function in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite element method (FEM) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions~\cite{Arroyo2006}. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)~\cite{Navas2018a} and Li {\it et al.} (2012)~\cite{Li2012} for
Optimal Transportation Meshfree (OTM) method. And more
recently under \acrshort{mpm} framework by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020} but without exploring the benefits of
the regularization parameter \gls{LME-beta}.

This techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using
techniques~\cite{Tran2019e}. These numerical inaccuracies also known
as wiggles are due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises is to add a
nonphysical damping source to the equilibrium equations, this
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing the accuracy of the time integration
scheme. One of the most popular is the \acrfull{igimp}
Charlton {\it et al.} (2017)~\cite{Charlton2017}, more recently Tran \&
Solowski (2019)~\cite{Tran2019e} proposed a generalised-$\alpha$ scheme
for \acrshort{mpm} with promising results but at the expense of increasing
the computational effort. In this paper a less time consuming and high
efficient explicit predictor corrector integration method has been
proposed. It consists in an accommodation of the Newmark
predictor-corrector or central difference explicit (CD). We have
choose this method among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because it simplicity and it good
performance dealing with solids dynamics problems under a mesh-free
framework in~\cite{Navas2018a}.

@@

The article is organized as follows. Section \ref{sec:Notation}
briefly reviews the notation here employed. Next, Section \ref{sec:derivation-mpm}
is devoted to present briefly the governing equations of the elastic
problem, the variational formulation and the Galerkin procedure. In Section \ref{sec:epc-algor-mpm} an explicit predictor-corrector
time integration scheme for \acrshort{mpm} is proposed. Section
\ref{sec:local-max-ent} briefly reviews the local \textit{max-ent}
basis functions here employed. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are presented. Finally,
conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{The meshfreesubsection procedure}
\label{sec:derivation-mpm}

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. In consequence it
is structured as follows: the governing equations will be introduced
in \ref{sec:The boundary-value-problem}, later the variational statement of
the problem will be presented \ref{sec:variational-formulation}, and
finally a discretization procedure thorough Galerkin is performed in
\ref{sec:Galerkin-procedure}. The method has three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation thorough a FEM-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carries on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstain subscripts
convention is adopted therefore repeated index means addition.

\subsection{Derivation of MPM procedure}
\label{sec:derivation-mpm}

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from \gls{u} the strain field \gls{strain} is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
\gls{stress}, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

To write the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative} represents the weak form
formulation of the problem.

@@

FEM, in \acrshort{mpm} a double discretization procedure is performed as we will

@@

\subsection{Explicit predictor-corrector scheme for \acrshort{mpm}.}
\label{sec:epc-algor-mpm}

In this section, a explicit predictor-corrector time integration
scheme is presented. It is based in the Newmark a-form 
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of the type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}
As \acrshort{mpm} has a nodal stage, it is possible to apply this methods
successfully in \acrshort{mpm} framework as was proved by
\cite{Tran2019e}. Taking the predictor definition from and calculating
nodal velocity, and updating particles position employing nodal values
of velocity and acceleration. The predictor-corrector algorithm has
been described in the classic literature \cite{Hughes2000}, and its
computational advantages and stability were widely proof by Liu
\cite{Xiaojian94}. The ``classic'' PCE algorithm starts with a
predicted value of the nodal velocity at the $(k+1)$th time step denoted by $\vec{\tilde{v}}_I^{k+1}$ given as,
\begin{equation}
  \label{eq:Predictor-velocity-I}
  \vec{\tilde{v}}_I^{k+1} = \vec{v}_I^k + (1 - \gamma)\ \Delta t\ \vec{a}_I^k
\end{equation}
In \eqref{eq:Predictor-velocity-I} arise a \textit{user-defined}
parameter $\gamma \geq 0$. This parameter influences both the predictor accuracy
and the stability of the algorithm. As point out Liu
\cite{Xiaojian94}, the truncation error of the predictor formula is
$O(\Delta t^3)$ when $\gamma = 0.5$, and is unconditionally stable if
$ 0 < \gamma \leq 0.25$.
